### 2018-08-05 contest/eda/eda.Rmd
Started some EDA.  At least Fed, Nadal, and SWilliams are in point by point.  We may want to use ATP/WTA match data otherwise

---
title: "EDA"
author: "The Competent Ones"
date: "August 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(devtools)
library(deuce)
library(knitr)
devtools::load_all("..") # Loads courtsports package
```

```{r data-extraction	}
data("point_by_point")
pbp = tbl_df(point_by_point)
pbp$wim = grepl("Wimbledon", pbp$tny_name)
pbp$french = grepl("FrenchOpen", pbp$tny_name)
pbp$aus = grepl("AustralianOpen", pbp$tny_name)
pbp$us = grepl("USOpen", pbp$tny_name)
pbp$qual = grepl("Qualifying", pbp$tny_name)

# Create data for Grand Slam finals
gsf = filter(pbp, !qual & (wim | french | us | aus))
```
```{r numbers}
dim(gsf)
colnames(gsf)
kable(table(gsf$tny_name))

any(grepl("Serena", gsf$server1))
any(grepl("Nadal", gsf$server1))
any(grepl("Federer", gsf$server1))

## Opponents of our 3
## Serena
gsf$server2[grepl("Serena", gsf$server1)]
## Nadal
gsf$server2[grepl("Nadal", gsf$server1)]
## Fed
gsf$server2[grepl("Federer", gsf$server1)]
```