### 2018-08-17 contest/eda/eda.Rmd
made some hsitograms

---
title: "EDA"
author: "The Competent Ones"
date: "August 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(devtools)
library(deuce)
library(knitr)
devtools::load_all("..") # Loads courtsports package
```

```{r data-extraction	}
data("point_by_point")
pbp = tbl_df(point_by_point)
pbp$wim = grepl("Wimbledon", pbp$tny_name)
pbp$french = grepl("FrenchOpen", pbp$tny_name)
pbp$aus = grepl("AustralianOpen", pbp$tny_name)
pbp$us = grepl("USOpen", pbp$tny_name)
pbp$qual = grepl("Qualifying", pbp$tny_name)

# Create data for Grand Slam finals
gsf = filter(pbp, !qual & (wim | french | us | aus))
```
```{r numbers}
dim(gsf)
colnames(gsf)
kable(table(gsf$tny_name))

any(grepl("Serena", gsf$server1))
any(grepl("Nadal", gsf$server1))
any(grepl("Federer", gsf$server1))

## Opponents of our 3
## Serena
gsf$server2[grepl("Serena", gsf$server1)]
## Nadal
gsf$server2[grepl("Nadal", gsf$server1)]
## Fed
gsf$server2[grepl("Federer", gsf$server1)]
```

## Looking at ATP (Men's) matches

```{r}
data(atp_matches)

gs <- atp_matches %>% filter(tourney_level == "Grand Slams" & year >= 2013)
sort(unique(gs$tourney_name))
colnames(gs)

## Add new column of points won
gs <- gs %>% mutate(w_pointswon = w_1stWon + w_2ndWon + l_svpt - l_1stWon - l_2ndWon)
gs <- gs %>% mutate(l_pointswon = l_1stWon + l_2ndWon + w_svpt - w_1stWon - w_2ndWon)
gs <- gs %>% mutate(w_morepts = w_pointswon >= l_pointswon)
n_correct <- sum(gs$w_morepts, na.rm = TRUE)
na_pt_indices <- which(is.na(gs$w_morepts))
n_correct / (nrow(gs) - length(na_pt_indices)) # more points won tells us who won >95% of the time
gs[na_pt_indices,]  ## All NA are straight up retirements so not really NA
incorrect_inds <- which(!gs$w_morepts)
table(gs[incorrect_inds, "tourney_name"])
table(gs[incorrect_inds, "Retirement"])

```

Australian Open with most wins with loser winning more points than winner.	Of all the matches where the winners won fewer points than the losers, only 10 were retirements.  Which means you are more likely to quit when you are down (10 night before).

```{r data-cleaning}
library(forcats)
library(stringr)
## Combine US Open
gs$tournament <- factor(gs$tourney_name)
gs$tournament <- gs$tournament %>% fct_collapse("US Open" = c("US Open", "Us Open"))

## Extract te number of sets in each match
no_ret <- gsub(" RET", "", gs$score)
set_scores <- str_split(no_ret, " ")
n_sets <- sapply(set_scores, length)
gs$n_sets <- n_sets

```
```{r some-plots}
library(ggplot2)
library(reshape2)

g <- ggplot(data = gs, aes(x = w_pointswon)) + geom_histogram()
g

g2 <- ggplot(data = gs, aes(x = w_pointswon)) +
    geom_histogram(aes(y=..density..)) +
    facet_grid(~tournament)
g2

##
bp <- ggplot(data = gs, aes(x = n_sets)) +
    geom_bar(aes(y = ..prop..)) +
    facet_grid(~tournament)
bp

g3 <- ggplot(data = gs, aes(x = w_pointswon)) +
    geom_histogram(aes(y=..density..)) +
    facet_grid(n_sets~tournament)
g3

```