### 2018-09-09 contest/surface-type/paper-tennis.tex
started intro and expanded data section.  Updated README

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Analyzing the Effect of Surface in Tennis Grand Slams},
            pdfauthor={Kayla Frisoli, Shannon Gallagher, and Amanda Luby},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Analyzing the Effect of Surface in Tennis Grand Slams}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Kayla Frisoli, Shannon Gallagher, and Amanda Luby}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{September 08, 2018}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}

\begin{document}
\maketitle
\begin{abstract}
Tennis grand slams consist of the Australian Open, French Open,
Wimbledon, and US Open, which are played on hard (Plexicushion), clay,
grass, and hard (DecoTurf) courts, respectively. The surface type may
substantially impact ball speed, height, and spin as well as player
speed and agility. It is also believed that play style and practice
habits may contribute to different results across surface types. For
example, Rafael Nadal is thought to be the best clay court player of all
time whereas Roger Federer is particularly known for dominance at
Wimbledon. On the women's side, Serena Williams once struggled on clay
courts but has seemingly transformed her style to perform better on clay
courts, but has perhaps suffered on grass as a consequence. In this
analysis, we examine the result of the top 100 players in grand slams
from 2013-2017 across the four different surfaces. We create a
hierarchical model with fixed and random effects to predict the number
of points won in a match. We take into consideration player-specific
effects, nationality (which is thought to have an effect on play style),
sex, ranking, ELO, and game statistics. We assess the fit of our model
using standard statistical techniques (e.g.~MSE, AIC, BIC, residual
diagnostics) in addition to `common knowledge' factors (for instance,
Rafael Nadal should be indicated as a superior clay court player by the
model). We compare the results of top 100 players across grand slams to
examine the effect of court surface. We also provide an in-depth
analysis of Nadal, Federer, and S. Williams.
\end{abstract}

\hypertarget{sec:iintro}{%
\section{Introduction}\label{sec:iintro}}

Rafael Nadal is known as the ``King of Clay'' in tennis, having won 11
out of his current 17 grand slams titles at the French Open, which is
played on a clay surface {[}cite{]}. In contrast, his rival Roger
Federer has won his most grand slam titles (8 out of 20) at Wimbledon,
which is played on grass. On the women's side, Serena Williams has been
dominant both on hard court (7 titles at the Australian Open and 6 at US
Open) and grass (7 at Wimbledon). This trend extends to other top
players, who seem to have better results at some grand slams than
others. More broadly, it seems that country of origin has an interaction
effect with court type. For example, Spanish players seem to excel on
clay courts and Americans have great success at Wimbledon despite grass
courts not being of wide use in the USA. It also worth questioning
whether the US Open and Australian Open should be grouped together as
hard courts despite having different surface compositions {[}cite{]}. In
this paper, we analyze the results of grand slam players from 2013-2017,
and we

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Determine if and how court surface effects players by implementation
  of a series of nested hierarchical models
\item
  Examine how Nadal, Federer, and Williams' play differs by surface
\item
  Assess whether we can group the two hard court surfaces together.
\end{enumerate}

As to issue (1) the hiearchical modelling, we use models similar to
{[}cite{]} and {[}cite{]}. More specifically, we model the players'
expected points in a match based on the player's own characteristics,
the court/tournament effects, and the oponent's ranking. We use a
Bayesian method because in {[}cite,cite{]} they had success in similar
situations.

For issue (2) the player analysis of Nadal, Federer, and Williams, we
examine whether our model passes the ``common sense''" tests (e.g.~these
players should win often when contrasting them to opponents) like in
{[}ventura something{]} and whether these players do have surface
apparent effects.

Finally, for issue (3), we use clustering methods {[}cite{]} in order to
determine which court surface types are more similar to one another.

Readers may object that we are looking at difference between grand
slams, which each have their own time period, weather conditions, play
time conditions, and ``home court effects'' instead of differences in
surfaces alone. However, (1) grand slam data is the most readily
available and most complete which makes it the best choice for
modelling, (2) we adjust for these confounders where we can, and (3)
analyzing the difference in the grand slams is still useful as they are
considered to be the most prestigious events in tennis.

The rest of this paper is organized as follows. In
\protect\hyperlink{sec:data}{Section Data} we describe our grand slam
tennis data. In \protect\hyperlink{sec:eda}{Section Early Data Analysis}
we examine the data at a high level and use clustering whether to
determine how the courts differ from one another. In
\protect\hyperlink{sec:methods}{Section Methods} we describe our
hierarchical models we use to determine difference in court surfaces. In
\protect\hyperlink{sec:results}{Section Results} we describe the results
of our modelling and also examine the play of Nadal, Federer, and
Williams. Finally in \protect\hyperlink{discussion}{Section Discussion},
we discuss future work and extensions or our model.

\hypertarget{sec:data-eda}{%
\section{Data and EDA}\label{sec:data-eda}}

\hypertarget{sec:data}{%
\subsection{Data}\label{sec:data}}

The data consists of 5080 matches split evenly over the four grand slams
and the two leagues: ATP (men's) and WTA (women's). Each match has 80
attributes, many of which are redundant. We focus on the following
attributes for both the winner and loser of the match: games won, points
won, retirement, break points faced, break points saved, aces, country
of origin, and player attributes. Additionally, we take into account the
number of sets in a match, the surface type, minutes played, and round
of the tournament. A subset of the data is shown in Table
\ref{tab:data}.

\rowcolors{2}{gray!6}{white}
\begin{table}

\caption{\label{tab:tab-data}\label{tab:data}Example of the grand slam data.  It includes winner and loser attributes, match attributes, and tournament attributes.  Not all attributes are shown here.}
\centering
\begin{tabular}[t]{lrlrlrrl}
\hiderowcolors
\toprule
Tournament & Year & W. Country & W. Points & Winner & W. Rank & L. Points & Loser\\
\midrule
\showrowcolors
US Open & 2014 & USA & 58 & Serena Williams & 1 & 31 & Taylor Townsend\\
US Open & 2013 & ESP & 136 & Rafael Nadal & 2 & 112 & Philipp Kohlschreiber\\
Australian Open & 2015 & ESP & 91 & Rafael Nadal & 3 & 51 & Mikhail Youzhny\\
US Open & 2013 & USA & 65 & Serena Williams & 1 & 41 & Yaroslava Shvedova\\
Australian Open & 2017 & USA & 73 & Serena Williams & 2 & 60 & Lucie Safarova\\
\bottomrule
\end{tabular}
\end{table}
\rowcolors{2}{white}{white}

The data is obtained from Jeff Sackman's open website {[}cite{]} via the
\texttt{R} package \texttt{deuce} {[}cite{]}. All steps of our analysis
from collection to dissemination are freely available online and are
compiled into an \texttt{R} package
\url{http://github.com/shannong19/courtsports}. The paper itself may be
generated from the file
\url{https://github.com/shannong19/courtsports/blob/master/contest/surface-type/paper-tennis.Rmd}.

\hypertarget{sec:eda}{%
\subsection{Early Data Analysis}\label{sec:eda}}

\hypertarget{sec:methods}{%
\section{Methods}\label{sec:methods}}

\hypertarget{sec:results}{%
\section{Results}\label{sec:results}}

\hypertarget{sec:discussion}{%
\section{Discussion}\label{sec:discussion}}

\hypertarget{sec:refs}{%
\section{References}\label{sec:refs}}

\end{document}