### 2018-09-09 contest/graphics/graphics.Rmd
add graphics

---
title: "Graphics"
author: "Kayla Frisoli, Shannon Gallagher, and Amanda Luby"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
library(deuce)
library(dplyr)
library(ggplot2)
library(forcats)
library(stringr)
library(knitr)
library(kableExtra)

## To change into library(courtsports)
devtools::load_all("../..")
```

```{r data-cleaning, include = FALSE}
## Load WTA and ATP matches from deuce package
data(atp_matches)
data(wta_matches)

## Clean data
gs <- clean_matches_data(atp_matches = atp_matches,
                         wta_matches = wta_matches,
                         start_year = 2013,
                         end_year = 2017,
                         include_quals = FALSE,
                         level = "Grand Slams")

## Extract sets, points, and games
gs <- extract_sgp(gs)

## Make into player data
gs_players <- matches_to_player_data(gs,
                                     w_opp_vars = "loser_rank",
                                     l_opp_vars = "winner_rank",
                                     opp_var_names = "opponent_rank")

```