### 2018-09-10 contest/graphics/graphics.Rmd
Merging

---
title: "Graphics"
author: "Kayla Frisoli, Shannon Gallagher, and Amanda Luby"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = TRUE)
library(deuce)
library(dplyr)
library(ggplot2)
library(forcats)
library(stringr)
library(knitr)
library(kableExtra)
library(pander)

## To change into library(courtsports)
devtools::load_all("../..")
```

```{r data-cleaning, include = FALSE}
## Load WTA and ATP matches from deuce package
data(atp_matches)
data(wta_matches)

## Clean data
gs <- clean_matches_data(atp_matches = atp_matches,
                         wta_matches = wta_matches,
                         start_year = 2013,
                         end_year = 2017,
                         include_quals = FALSE,
                         level = "Grand Slams")

## Extract sets, points, and games
gs <- extract_sgp(gs)

## Make into player data
gs_players <- matches_to_player_data(gs,
                                     w_opp_vars = "loser_rank",
                                     l_opp_vars = "winner_rank",
                                     opp_var_names = "opponent_rank")

```

## Is being a winner from Spain independent of tournament?

```{r, fig.height=5, fig.width=8, fig.align='center'}

gs <- mutate(gs, 
             winner_ioc_spain = ifelse(winner_ioc == "ESP", "Spain", "Other"))

props <- gs %>% group_by(tournament, winner_ioc_spain) %>%
        tally() %>%
        group_by(tournament) %>%
        mutate(pct = n / sum(n))

ggplot(props, aes(tournament, pct, fill = winner_ioc_spain)) +
        geom_bar(stat = "identity") +
        scale_fill_manual("Country of origin",
                          values = c( "#FFC400", "#C60B1E")) + 
  labs(x = " ", y = "Proportion", title = "Spanish players win the French Open more often than other tournaments") + 
  geom_text(data = filter(props, winner_ioc_spain == "Spain"), 
            aes(tournament, y = pct + .05, label = paste0(round(pct, 2),"%")),
                       size=4)

```

```{r sample, echo=FALSE, results='asis'}
library(xtable)

spain_vs_tourn <- with(gs, table(tournament, winner_ioc_spain))

t1 <- kable(spain_vs_tourn, format = "latex", booktabs = TRUE)
t2 <- kable(chisq.test(spain_vs_tourn)$expected, format = "latex", booktabs = TRUE)

cat(c("\\begin{table}[!htb]
    \\begin{minipage}{.5\\linewidth}
      \\caption{Actual table}
      \\centering",
        t1,
    "\\end{minipage}%
    \\begin{minipage}{.5\\linewidth}
      \\centering
        \\caption{Expected table}",
        t2,
    "\\end{minipage} 
\\end{table}"
))  
```

```{r}
chisq.test(spain_vs_tourn) %>% 
  pander()
```