### 2018-09-11 contest/model-fitting/model-comparisons.Rmd
please work

---
title: "Model Comparisons"
author: "Amanda Luby, Shannon Gallagher, Kayla Frisoli"
date: "9/10/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(dplyr)
load("stan-model-fits.RData")
load(file = "../../data/gs_players.rda")
load("../../data/gs.rda")

gs_players$name_int = as.integer(as.factor(gs_players$name))
gs_players$hand_fac = as.factor(gs_players$hand)
gs_players$ioc_fac = as.factor(gs_players$ioc)
gs_players$atp = gs_players$league == "ATP"
gs_players$year_fac = as.factor(gs_players$year)
```

### Model Fit

* Took minutes out because it was disproportionately missing for women players (and missing data gets thrown out)
* Individual effects for tournament don't really seem to make a difference - mostly wrapped up in other variables

```{r}
linmod = lm(pointswon ~ ioc_fac +  tournament + round + rank + opponent_rank + year_fac + atp, data = gs_players)
country_effects = lmer(pointswon ~ round + rank + opponent_rank + year_fac + atp + (0 + tournament |ioc_fac), data = gs_players)
ind_effects = lmer(pointswon ~  ioc_fac + round + rank + opponent_rank + year_fac + atp + (0 + tournament |name_int), data = gs_players)

par(mfrow=c(1,3))
plot(residuals(linmod)~fitted(linmod), pch =19, cex = .5, main = "Model 1",
     xlab = "Fitted", ylab = "Residuals")
plot(residuals(country_effects) ~ fitted(country_effects), pch =19, cex = .5, main = "Model 2",
     xlab = "Fitted", ylab = "Residuals")
plot(residuals(ind_effects) ~ fitted(ind_effects), pch =19, cex = .5, main = "Model 3",
     xlab = "Fitted", ylab = "Residuals")

```

```{r}
extractAIC(linmod)
extractAIC(country_effects)
extractAIC(ind_effects)
```

### Model Comparison (tennis knowledge)

```{r}
superstars = gs_players %>% 
  filter(., name == "Serena Williams" | name == "Rafael Nadal" | name == "Roger Federer")

par(mfrow = c(1,3))
plot(predict(linmod, newdata = superstars) ~ superstars$pointswon, col = as.factor(superstars$name))
plot(predict(country_effects, newdata = superstars) ~ superstars$pointswon, col = as.factor(superstars$name))
plot(predict(ind_effects, newdata = superstars) ~ superstars$pointswon, col = as.factor(superstars$name))
```

Sanity Checks:

* Serena should be good on hard surface
* Nadal should be good on clay 
* Federer should be good on grass

```{r}

```