### 2018-09-14 contest/model-fitting/model-comparisons.Rmd

---
title: "Model Comparisons"
author: "Amanda Luby, Shannon Gallagher, Kayla Frisoli"
date: "9/10/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(dplyr)
library(kableExtra)
#load("stan-model-fits.RData")
load(file = "../../data/gs_players.rda")
load("../../data/gs.rda")

gs_players$name_int = as.integer(as.factor(gs_players$name))
gs_players$hand_fac = as.factor(gs_players$hand)
gs_players$ioc_fac = as.factor(gs_players$ioc)
gs_players$atp = gs_players$league == "ATP"
gs_players$year_fac = as.factor(gs_players$year)
gs_players$late_round = gs_players$round >= "R16"

```

### Model Fit

* Took minutes out because it was disproportionately missing for women players (and missing data gets thrown out)
* Initially began by modeling points won in the match, but found that this was not the best approach because more points won does not necessarily correspond to a better performance.

Variables considered as predictors:

* `ioc_fac`: Country that the player represents
* `tournament`: Which Grand Slam the match was played at
* `late_round`: Indicator for whether the match occured in Round of 16 or later
* `rank`: Rank of player at the time of the match
    + Included on log scale 
* `opponent_rank`: Rank of opponent at the time of the match
    + Included on log scale
* `year`: Factor variable with a level for each year included in the dataset (2013-2017)
* `atp`: Indicator that the match was played in the ATP league instead of the WTF league

Models considered:

* No mixed effects
    + `nocourt_logistic`: $\text{logit}(\pi_i) = \beta X$, $X$ does not include `tournament`
        + If this model fits well, court surface does not have an effect on winning probability
    + `base_logistic`: $\text{logit}(\pi_i) = \beta X$
        + If this model fits better than the mixed-effects models, performance does not depend on individual-level or IOC-level 
* IOC-level effect
    + `country_logistic`: $\text{logit}(\pi_i) = \beta X + \beta_{T[i]}T$, $beta_{T[i]} = \alpha_i C_i$

```{r}
nocourt_logistic = glm(did_win ~ ioc_fac + late_round + log(rank) + log(opponent_rank) + year_fac + atp, data = gs_players, family = "binomial")
base_logistic = glm(did_win ~ ioc_fac +  tournament + late_round + log(rank) + log(opponent_rank) + year_fac + atp, data = gs_players, family = "binomial")
country_logistic = glmer(did_win ~ late_round + log(rank) + log(opponent_rank) + year_fac + atp + (0 + tournament |ioc_fac), data = gs_players, family = "binomial")
ind_logistic = glmer(did_win ~  ioc_fac + late_round + log(rank) + log(opponent_rank) + year_fac + atp + (0 + tournament |name_int), data = gs_players, family = "binomial")
ind_logistic_noioc = glmer(did_win ~ late_round + log(rank) + log(opponent_rank) + year_fac + atp + (0 + tournament |name_int), data = gs_players, family = "binomial")
ind_int_logistic = glmer(did_win ~ late_round + log(rank) + log(opponent_rank) + year_fac + atp + tournament + (1|name_int), data = gs_players, family = "binomial")
ind_year_logistic = glmer(did_win ~ late_round + log(rank) + log(opponent_rank) + atp + tournament + (0+year_fac|name_int), data = gs_players, family = "binomial")

model_names = c("nocourt_logistic", "base_logistic", "country_logistic", "ind_logistic", 
                "ind_logistic_noioc", "ind_int_logistic", "ind_year_logistic")
model.sums = data.frame(model = model_names, aic = rep(NA, length(model_names)), edf = rep(NA,length(model_names)))

for(mod in 1:nrow(model.sums)){
  name = model_names[mod]
  model.sums$aic[mod] = extractAIC(get(name))[2]
  model.sums$edf[mod] = extractAIC(get(name))[1]
}

kable(model.sums)
```

Based on AIC, the `ind_logistic_noioc` model performed best after accounting for the number of variables in the model.

### Model Comparison (tennis knowledge)

```{r}
superstars = gs_players %>% 
  filter(., name == "Serena Williams" | name == "Rafael Nadal" | name == "Roger Federer")

par(mfrow = c(1,3))
plot(predict(linmod, newdata = superstars) ~ superstars$pointswon, col = as.factor(superstars$name))
plot(predict(country_effects, newdata = superstars) ~ superstars$pointswon, col = as.factor(superstars$name))
plot(predict(ind_effects, newdata = superstars) ~ superstars$pointswon, col = as.factor(superstars$name))
```

Sanity Checks:

* Serena should be good on hard surface
* Nadal should be good on clay 
* Federer should be good on grass