### 2018-09-14 contest/graphics/graphics.Rmd
update grpahics file; eval=FALSE some stuff

---
title: "Graphics"
author: "We have no names"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(cache = TRUE)
library(deuce)
library(dplyr)
library(ggplot2)
library(forcats)
library(stringr)
library(knitr)
library(kableExtra)
library(pander)
library(gridExtra)
library(grid)

## To change into library(courtsports)
devtools::load_all("../..")
```

```{r data-cleaning, include = FALSE}
data("gs")
data("gs_players")
data("gs_partial_pbp")

my_theme <-  theme_bw() + # White background, black and white theme
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 14,
        family="serif"),
        plot.title = element_text(hjust = 0.5, size=16))

tournament_colors <- c("#0297DB", "#b06835", "#0C2340", "#54008b")

league_colors <- c("#0C2340", "#902BA3")

win_colors <- c("#336699", "#339966")

```

## Examining the distribution of points earned

We first examine the distibution of points earned per match. We find that the distribution is similar across tournament, with Wimbledon differing slightly from the other grand slams. As expected, there are more points earned in the WTA than the ATP due to the differing numbers of games played. Also unsuprisingly, the winners of the match tended to earn more points than the losers.

```{r, fig.height=5, fig.width=12}

# grid.arrange(p1, p2, p3, ncol=3, widths=c(.25, .41, .34))

p2 <- ggplot(gs_players) + geom_density(aes(pointswon, color = tournament)) + 
  labs(x = "Points") + my_theme +
  scale_color_manual("Tournament", values=tournament_colors)

p3 <- ggplot(gs_players) + geom_density(aes(pointswon, color = league))+ 
  labs(x = "Points") +
  scale_color_manual("League", values=league_colors) +
  my_theme

# grid.arrange( p2, p3, ncol=2,
#              top = textGrob("Winning points by tournament and league",
#                             gp=gpar(fontsize=20)))

p1 <- ggplot(gs_players) + geom_density(aes(pointswon, color = factor(did_win))) + 
  labs(x = "Points") +
  scale_color_manual("Winner?", values=win_colors) +
  my_theme +
  facet_wrap(~league, ncol=1)

grid.arrange(p1, p2, ncol=2,
             top = textGrob("Distribution of points per match",
                            gp=gpar(fontsize=20, fontfamily="serif")),
              layout_matrix = rbind(c(NA, 1),
                                    c(2, 1),
                                    c(NA, 1)),
             heights=c(1, 3, 1))

```

## Home court advantage

It is commonly thought that there is a home court advantage in grand slam games (SOURCE). In our data we find this to be true (i.e. French players win the French open more than French players win other slams). But, we also know that the home team is given preference for wild card bids (SOURCE) so potentially citizens of a particular country play in "their" tournament more often than they play in other tournaments. We also find this to be true in our data for France, The United States, and Australia (i.e. the proportion of French players in the French Open is greater than the proportion of French players in other slams).

Therefore, we want to see how the proportion of wins for the home country changes across the different tournaments. If there was really a home court advantage, the proportion of French wins each round would be higher at the French Open than the Australian Open, the US Open, and Wimbledon. The same would be true for Australia and the US. But, we see that this isn't the case. After accounting for the number of players from each country, we don't find a home court advantage in the grand slam.

```{r, fig.height=4, fig.width=12}
france_wins <- filter(gs_players, ioc == "FRA", round < "QF") %>% 
  group_by(tournament, round) %>% 
  summarise(win_prop = mean(did_win),
            win_prop_se = sqrt(win_prop*(1-win_prop)/n()))

us_wins <- filter(gs_players, ioc == "USA", round < "QF") %>% 
  group_by(tournament, round) %>% 
  summarise(win_prop = mean(did_win),
            win_prop_se = sqrt(win_prop*(1-win_prop)/n()))

aus_wins <- filter(gs_players, ioc == "AUS", round < "QF") %>% 
  group_by(tournament, round) %>% 
  summarise(win_prop = mean(did_win),
            win_prop_se = sqrt(win_prop*(1-win_prop)/n()))

p1<- ggplot(france_wins, aes(round, win_prop, color=tournament, group=tournament)) + 
  geom_line() + 
  # geom_point() +
  labs(title = "French players",
       x = "", y = "Win proportion") + 
  scale_color_manual(guide=FALSE, values=tournament_colors) + my_theme

p2 <- ggplot(us_wins, aes(round, win_prop, color=tournament, group=tournament)) + 
  geom_line() + 
  # geom_point() +
  labs(title = "American players",
       x = "", y = "Win proportion") +  
  scale_color_manual(guide=FALSE, values=tournament_colors)  + my_theme

p3 <- ggplot(aus_wins, aes(round, win_prop, color=tournament, group=tournament)) + 
  geom_line() + 
  # geom_point() +
  labs(title = "Australian players",
       x = "", y = "Win proportion") + 
  scale_color_manual(" ", values=tournament_colors) + my_theme

grid.arrange(p1,p2, p3, nrow=1, widths = c(.29, .29, .42),
             top = textGrob("Home court advantage debunked: win proportion by country",
                            gp=gpar(fontsize=20)),
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           # x = 0,
                           # y = 0.5,
                           just = "left"))

```

## Spaniards on clay

```{r}
filter(gs_players, round == "R128", ioc == "ESP") %>% 
  group_by(tournament) %>% 
  summarise(avg_ranking = mean(rank))

spain_ranks <- filter(gs_players, ioc == "ESP") %>% 
  group_by(tournament, round) %>% 
  summarise(avg_ranking = mean(rank))

ggplot(spain_ranks) + 
  geom_line(aes(round, avg_ranking, color=tournament, group=tournament)) + 
  labs(title = "Average ranking among Spaniards")

```

# Points won

```{r, fig.height=4, fig.width=12}
p2 <- ggplot(gs_players) + geom_density(aes(pointswon, color = tournament)) + 
  labs(x = "Winner points") + my_theme +
  scale_color_manual("Tournament", values=tournament_colors)

p3 <- ggplot(gs_players) + geom_density(aes(pointswon, color = league))+ 
  labs(x = "Winner points") +
  scale_color_manual("League", values=league_colors) +
  my_theme

# grid.arrange( p2, p3, ncol=2,
#              top = textGrob("Winning points by tournament and league",
#                             gp=gpar(fontsize=20)))
```

```{r, fig.height=3.5, fig.width=10, fig.align='center'}

p1 <- ggplot(gs) + geom_point(aes(w_pointswon, l_pointswon)) + 
  geom_abline(slope=1, intercept=0, color="red") + 
  labs(x = "Winner points", y = "Loser points")

p2 <- ggplot(gs) + geom_point(aes(w_pointswon, l_pointswon, color = league), alpha=.3) + 
    scale_color_manual(values=c("#339966", "#336699")) +
  geom_abline(slope=1, intercept=0, color="red") + 
  geom_smooth(aes(w_pointswon, l_pointswon, color = league), method='lm') + 
  labs(x = "Winner points", y = "Loser points")

grid.arrange(p1, p2, ncol=2, widths=c(.48, .52))

```

```{r, fig.height=4, fig.width=12}
ggplot(filter(gs_players, league == "WTA", round == "R128")) + 
  geom_point(aes(rank, pointswon, color=factor(did_win)), alpha=.5) + 
  labs(x = "Rank", y = "Points", title="Round 128")

```

```{r, fig.height=4, fig.width=12}
ggplot(filter(gs)) + 
  geom_point(aes(loser_rank, winner_rank))

```

## Games won

```{r, fig.height=4, fig.width=12}

p1 <- ggplot(gs_players) + geom_histogram(aes(gameswon), bins=30)+ 
  labs(x = "Winner games")

p2 <- ggplot(gs_players) + geom_density(aes(gameswon, color = tournament))+ 
  labs(x = "Winner games")

p3 <- ggplot(gs_players) + geom_density(aes(gameswon, color = league))+ 
  labs(x = "Winner games")

grid.arrange(p1, p2, p3, ncol=3, widths=c(.25, .41, .34))

```

```{r, fig.height=3.5, fig.width=10, fig.align='center'}

p1 <- ggplot(gs) + geom_point(aes(w_gameswon, l_gameswon)) + 
  geom_abline(slope=1, intercept=0, color="red") + 
  labs(x = "Winner games", y = "Loser games")

p2 <- ggplot(gs) + geom_point(aes(w_gameswon, l_gameswon, color = league), alpha=.3) + 
    scale_color_manual(values=c("#339966", "#336699")) +
  geom_abline(slope=1, intercept=0, color="red") + 
  geom_smooth(aes(w_gameswon, l_gameswon, color = league), method='lm')  + 
  labs(x = "Winner games", y = "Loser games")

grid.arrange(p1, p2, ncol=2, widths=c(.48, .52))

```

# Spain vs. tournament

## Is being a winner from Spain independent of tournament?

```{r, fig.height=3.2, fig.width=6, fig.align='center'}
gs <- mutate(gs, 
             winner_ioc_spain = ifelse(winner_ioc == "ESP", "Spain", "Other"))

props <- gs %>% group_by(tournament, winner_ioc_spain) %>%
        tally() %>%
        group_by(tournament) %>%
        mutate(pct = n / sum(n))

ggplot(props, aes(tournament, pct, fill = winner_ioc_spain)) +
        geom_bar(stat = "identity") +
        scale_fill_manual("Country of origin",
                          values = c( "#FFC400", "#C60B1E")) + 
  labs(x = " ", y = "Proportion", title = "Spanish players win the French Open\nmore often than other tournaments") + 
  geom_text(data = filter(props, winner_ioc_spain == "Spain"), 
            aes(tournament, y = pct + .05, label = paste0(round(pct*100, 2),"%")),
                       size=4) + theme_bw() + theme(plot.title = element_text(hjust = 0.5))
```

```{r sample, echo=FALSE, results='asis'}
library(xtable)

spain_vs_tourn <- with(gs, table(tournament, winner_ioc_spain))

t1 <- kable(spain_vs_tourn, format = "latex", booktabs = TRUE)
t2 <- kable(chisq.test(spain_vs_tourn)$expected, format = "latex", booktabs = TRUE)
spain_vs_tourn_chisq <- chisq.test(spain_vs_tourn)

t1 <- kable(spain_vs_tourn, format = "latex", booktabs = TRUE)
t2 <- kable(spain_vs_tourn_chisq$expected, format = "latex", booktabs = TRUE)

cat(c("\\begin{table}[!htb]
    \\begin{minipage}{.5\\linewidth}
      \\caption{Actual table}
      \\centering",
        t1,
    "\\end{minipage}%
    \\begin{minipage}{.5\\linewidth}
      \\centering
        \\caption{Expected table}",
        t2,
    "\\end{minipage} 
\\end{table}"
))  
```

```{r}
spain_vs_tourn_chisq %>% 
  pander()
```

\newpage

## But do Spaniards enter the French open more than they do other tournaments?

We look at players in the *first round* of all tournaments.

```{r, fig.height=3.2, fig.width=6, fig.align='center'}

mini.d <- bind_rows(select(filter(gs, round == "R128"), tournament, winner_ioc) %>% rename(ioc = winner_ioc),
            select(filter(gs, round == "R128"), tournament, loser_ioc) %>% rename(ioc = loser_ioc)) %>%
  mutate(ioc_spain = ifelse(ioc == "ESP", "Spain", "Other"))

props <- mini.d %>% group_by(tournament, ioc_spain) %>%
        tally() %>%
        group_by(tournament) %>%
        mutate(pct = n / sum(n))

ggplot(props, aes(tournament, pct, fill = ioc_spain)) +
        geom_bar(stat = "identity") +
        scale_fill_manual("Country of origin",
                          values = c( "#FFC400", "#C60B1E")) + 
  labs(x = " ", y = "Proportion",
       title = "Spanish players enter the French Open\nmore often than other tournaments") + 
  geom_text(data = filter(props, ioc_spain == "Spain"), 
            aes(tournament, y = pct + .05, label = paste0(round(pct*100, 2),"%")),
                       size=4)  + theme_bw() + theme(plot.title = element_text(hjust = 0.5))

```

```{r}
with(mini.d, table(tournament, ioc_spain)) %>%
  chisq.test() %>%
  pander()
```

Yes, but not significantly more.

## Spain proportion of wins across the tournaments

```{r}
spain_wins <- filter(gs_players, ioc == "ESP") %>% 
  group_by(tournament, round) %>% 
  summarise(win_prop = mean(did_win),
            win_prop_se = sqrt(win_prop*(1-win_prop)/n()))

ggplot(spain_wins, aes(round, win_prop, color=tournament, group=tournament)) + 
  geom_line(cex=2) + 
  geom_point() +
  labs(title = "Proportion of wins among Spaniards") +
  geom_errorbar(aes(ymin = win_prop - 2*win_prop_se,
                    ymax = win_prop + 2*win_prop_se),
                width=.2,
                 position=position_dodge(0.05))

```

# Player rankings over time

```{r}

# filter(gs_players, round == "R128") %>%
#   select(tourney_start_date, rank, )
#   ggplot(aes(x = tourney_start_date, y = rank)) + geom_line(group = name)
#   
filter(gs_players, round == "R128") %>%
  rename(player_name = name) %>%
  ggplot(aes(x = tourney_start_date, y = rank, color=player_name)) + 
  geom_line(aes(group = factor(player_name))) + guides(color=FALSE, group=FALSE)

```

# Home court advantage

## France

```{r, fig.height=3.2, fig.width=6, fig.align='center'}

gs <- mutate(gs, 
             winner_ioc_france = ifelse(winner_ioc == "FRA", "France", "Other"))

props <- gs %>% group_by(tournament, winner_ioc_france) %>%
        tally() %>%
        group_by(tournament) %>%
        mutate(pct = n / sum(n))

ggplot(props, aes(tournament, pct, fill = factor(winner_ioc_france, levels = c("Other", "France")))) +
        geom_bar(stat = "identity") +
        scale_fill_manual("Country of origin",
                          values = c("#F42A41", "#00209F")) + 
  labs(x = " ", y = "Proportion", title = "French players win the French Open\nmore often than other tournaments") + 
  geom_text(data = filter(props, winner_ioc_france == "France"), 
            aes(tournament, y = pct + .05, label = paste0(round(pct*100, 2),"%")),
                       size=4)  + theme_bw() + theme(plot.title = element_text(hjust = 0.5))

```

```{r, echo=FALSE, results='asis'}
library(xtable)

france_vs_tourn <- with(gs, table(tournament, winner_ioc_france))

france_vs_tourn_chisq <- chisq.test(france_vs_tourn)

t1 <- kable(france_vs_tourn, format = "latex", booktabs = TRUE)
t2 <- kable(france_vs_tourn_chisq$expected, format = "latex", booktabs = TRUE)

cat(c("\\begin{table}[!htb]
    \\begin{minipage}{.5\\linewidth}
      \\caption{Actual table}
      \\centering",
        t1,
    "\\end{minipage}%
    \\begin{minipage}{.5\\linewidth}
      \\centering
        \\caption{Expected table}",
        t2,
    "\\end{minipage} 
\\end{table}"
))  
```

```{r}
france_vs_tourn_chisq %>% 
  pander()
```

```{r, fig.height=3.2, fig.width=6, fig.align='center'}

mini.d <- mini.d %>%
  mutate(ioc_france = ifelse(ioc == "FRA", "France", "Other"))

props <- mini.d %>% group_by(tournament, ioc_france) %>%
        tally() %>%
        group_by(tournament) %>%
        mutate(pct = n / sum(n))

ggplot(props, aes(tournament, pct, fill = factor(ioc_france, levels = c("Other", "France")))) +
        geom_bar(stat = "identity") +
        scale_fill_manual("Country of origin",
                          values = c("#F42A41", "#00209F")) + 
  labs(x = " ", y = "Proportion",
       title = "French players enter the French Open\nmore often than other tournaments") + 
  geom_text(data = filter(props, ioc_france == "France"), 
            aes(tournament, y = pct + .05, label = paste0(round(pct*100, 2),"%")),
                       size=4)  + theme_bw() + theme(plot.title = element_text(hjust = 0.5))

```

```{r}
with(mini.d, table(tournament, ioc_france)) %>%
  chisq.test() %>%
  pander()
```

*Yes, significantly.*

\newpage

## United States

```{r, fig.height=3.2, fig.width=6, fig.align='center'}

gs <- mutate(gs, 
             winner_ioc_usa = ifelse(winner_ioc == "USA", "USA", "Other"))

props <- gs %>% group_by(tournament, winner_ioc_usa) %>%
        tally() %>%
        group_by(tournament) %>%
        mutate(pct = n / sum(n))

ggplot(props, aes(tournament, pct, fill = factor(winner_ioc_usa, levels = c("Other", "USA")))) +
        geom_bar(stat = "identity") +
        scale_fill_manual("Country of origin",
                          values = c( "#002868", "#BF0A30")) + 
  labs(x = " ", y = "Proportion", title = "American players win the U.S. Open\nmore often than other tournaments") + 
  geom_text(data = filter(props, winner_ioc_usa == "USA"), 
            aes(tournament, y = pct + .05, label = paste0(round(pct*100, 2),"%")),
                       size=4,
            color="white")  + theme_bw() + theme(plot.title = element_text(hjust = 0.5))

```

```{r, echo=FALSE, results='asis'}
usa_vs_tourn <- with(gs, table(tournament, winner_ioc_usa))

usa_vs_tourn_chisq <- chisq.test(usa_vs_tourn)

t1 <- kable(usa_vs_tourn, format = "latex", booktabs = TRUE)
t2 <- kable(usa_vs_tourn_chisq$expected, format = "latex", booktabs = TRUE)

# cat(c("\\begin{table}[!htb]
#     \\begin{minipage}{.5\\linewidth}
#       \\caption{Actual table}
#       \\centering",
#         t1,
#     "\\end{minipage}%
#     \\begin{minipage}{.5\\linewidth}
#       \\centering
#         \\caption{Expected table}",
#         t2,
#     "\\end{minipage} 
# \\end{table}"
# ))  
```

```{r}
usa_vs_tourn_chisq %>% 
  pander()
```

```{r, fig.height=3.2, fig.width=6, fig.align='center'}

mini.d <- mini.d %>%
  mutate(ioc_usa = ifelse(ioc == "USA", "USA", "Other"))

props <- mini.d %>% group_by(tournament, ioc_usa) %>%
        tally() %>%
        group_by(tournament) %>%
        mutate(pct = n / sum(n))

ggplot(props, aes(tournament, pct, fill = factor(ioc_usa, levels = c("Other", "USA")))) +
        geom_bar(stat = "identity") +
        scale_fill_manual("Country of origin",
                          values = c( "#002868", "#BF0A30")) + 
  labs(x = " ", y = "Proportion",
       title = "American players enter the U.S. Open\nmore often than other tournaments") + 
  geom_text(data = filter(props, ioc_usa == "USA"), 
            aes(tournament, y = pct + .05, label = paste0(round(pct*100, 2),"%")),
                       size=4,
            color="white")  + theme_bw() + theme(plot.title = element_text(hjust = 0.5))

```

```{r}
with(mini.d, table(tournament, ioc_usa)) %>%
  chisq.test() %>%
  pander()
```

## Australia

```{r, fig.height=3.2, fig.width=6, fig.align='center'}

gs <- mutate(gs, 
             winner_ioc_Australia = ifelse(winner_ioc == "AUS", "Australia", "Other"))

props <- gs %>% group_by(tournament, winner_ioc_Australia) %>%
        tally() %>%
        group_by(tournament) %>%
        mutate(pct = n / sum(n))

ggplot(props, aes(tournament, pct, fill = factor(winner_ioc_Australia, levels = c("Other", "Australia")))) +
        geom_bar(stat = "identity", color="grey") +
        scale_fill_manual("Country of origin",
                          values = c("#FFFFFF", "#012169")) + 
  labs(x = " ", y = "Proportion", title = "Australian players win the Australian Open\nmore often than other tournaments") + 
  geom_text(data = filter(props, winner_ioc_Australia == "Australia"), 
            aes(tournament, y = pct + .05, label = paste0(round(pct*100, 2),"%")),
                       size=4)  + theme_bw() + theme(plot.title = element_text(hjust = 0.5))

```

```{r, echo=FALSE, results='asis'}
Australia_vs_tourn <- with(gs, table(tournament, winner_ioc_Australia))

Australia_vs_tourn_chisq <- chisq.test(Australia_vs_tourn)

t1 <- kable(Australia_vs_tourn, format = "latex", booktabs = TRUE)
t2 <- kable(Australia_vs_tourn_chisq$expected, format = "latex", booktabs = TRUE)

cat(c("\\begin{table}[!htb]
    \\begin{minipage}{.5\\linewidth}
      \\caption{Actual table}
      \\centering",
        t1,
    "\\end{minipage}%
    \\begin{minipage}{.5\\linewidth}
      \\centering
        \\caption{Expected table}",
        t2,
    "\\end{minipage} 
\\end{table}"
))  
```

```{r}
Australia_vs_tourn_chisq %>% 
  pander()
```

```{r, fig.height=3.2, fig.width=6, fig.align='center'}

mini.d <- mini.d %>%
  mutate(ioc_Australia = ifelse(ioc == "AUS", "Australia", "Other"))

props <- mini.d %>% group_by(tournament, ioc_Australia) %>%
        tally() %>%
        group_by(tournament) %>%
        mutate(pct = n / sum(n))

ggplot(props, aes(tournament, pct, fill = factor(ioc_Australia, levels = c("Other", "Australia")))) +
        geom_bar(stat = "identity", color="grey") +
        scale_fill_manual("Country of origin",
                          values = c("#FFFFFF", "#012169")) + 
  labs(x = " ", y = "Proportion",
       title = "Australian players enter the Australian Open\nmore often than other tournaments") + 
  geom_text(data = filter(props, ioc_Australia == "Australia"), 
            aes(tournament, y = pct + .05, label = paste0(round(pct*100, 2),"%")),
                       size=4)  + theme_bw() + theme(plot.title = element_text(hjust = 0.5))

```

```{r}
with(mini.d, table(tournament, ioc_Australia)) %>%
  chisq.test() %>%
  pander()
```

# Conclusions

* Spaniards win the French Open more often than they win other tournaments, and this difference is significant. They play in the French Open more often than they play in other tournaments, but not significantly.

* French win the French Open more often than they win other tournaments, but this difference is (barely) insignificant. They play in the French Open more often than they play in other tournaments, at a significantly higher rate.

* Americans win the US Open more often than they win other tournaments, and this difference is significant. But, they play in the US Open more often than they play in other tournaments, at a significantly higher rate.

* Australians win the US Open more often than they win other tournaments, and this difference is significant. But, they play in the Australian Open more often than they play in other tournaments, at a significantly higher rate.

# Pairs

```{r, fig.height=12, fig.width=12, eval=FALSE}
library(GGally)

vars <- c("ht", "age", "rank", "opponent_rank", "ace", "df", "svpt", "1stIn", "SvGms", "bpSaved",
          "bpFaced", "pointswon", "gameswon")

ggpairs(gs_players,
        columns = which(names(gs_players) %in% vars),
        mapping=ggplot2::aes(colour = factor(did_win)))

```

## Follow-up

```{r, fig.height=3, fig.width=4}
ggplot(gs_players) + geom_point(aes(log(ht+1), log(ace+1), color = factor(did_win)))
```

# MDS

```{r, eval=FALSE}
gs_num <- select(gs, w_ace, w_df, w_svpt, w_1stIn, w_1stWon, w_2ndWon,
                w_bpSaved, w_bpFaced) %>%
na.omit()

mds <- cmdscale(dist(scale(gs_num)), k = 2) %>%
  as.data.frame()  # default column names will be V1, V2

mds$tournament <- select(gs, w_ace, w_df, w_svpt, w_1stIn, w_1stWon, w_2ndWon,
                w_bpSaved, w_bpFaced, tournament) %>%
na.omit() %>% pull( tournament)

ggplot(mds, aes(V1, V2, label = tournament, color=tournament)) + geom_text()

gs_num <- select(gs, w_ace, w_df, w_svpt, w_1stIn, w_1stWon, w_2ndWon,
                w_bpSaved, w_bpFaced, winner_rank) %>%
na.omit()

mds <- cmdscale(dist(scale(gs_num)), k = 2) %>%
  as.data.frame()  # default column names will be V1, V2

mds$tournament <- select(gs, w_ace, w_df, w_svpt, w_1stIn, w_1stWon, w_2ndWon,
                w_bpSaved, w_bpFaced, winner_rank, tournament) %>%
na.omit() %>% pull( tournament)

ggplot(mds, aes(V1, V2, label = tournament, color=tournament)) + geom_text()

```

```{r, eval=FALSE}
gs_num <- select(gs, w_ace, w_df, w_svpt, w_1stIn, w_1stWon, w_2ndWon,
                w_bpSaved, w_bpFaced) %>%
na.omit()

hclust_results <- hclust(dist(scale(gs_num)))

# plot(hclust_results)

colors_to_use <- select(gs, w_ace, w_df, w_svpt, w_1stIn, w_1stWon, w_2ndWon,
                w_bpSaved, w_bpFaced, tournament) %>%
  na.omit() %>% pull(tournament) %>% as.numeric()

library(dendextend)
dend <- as.dendrogram(hclust_results)
colors_to_use <- colors_to_use[order.dendrogram(dend)]

labels_colors(dend) <- colors_to_use
plot(dend, main = "A color for every tournament")

mds$tournament <- select(gs, w_ace, w_df, w_svpt, w_1stIn, w_1stWon, w_2ndWon,
                w_bpSaved, w_bpFaced, tournament) %>%
na.omit() %>% pull( tournament)

ggplot(mds, aes(V1, V2, label = tournament, color=tournament)) + geom_text()
```

```{r, eval=FALSE}
gs_num <- select(gs, w_ace, w_df, w_svpt, w_1stIn,
                w_bpSaved, w_bpFaced) %>%
na.omit()

mds <- cmdscale(dist(scale(gs_num)), k = 2) %>%
  as.data.frame()  # default column names will be V1, V2

mds$tournament <- select(gs, w_ace, w_df, w_svpt, w_1stIn,
                w_bpSaved, w_bpFaced, tournament) %>%
na.omit() %>% pull( tournament)

ggplot(mds, aes(V1, V2, label = tournament, color=tournament)) + geom_text()
```

# MDS

```{r, eval=FALSE}
# winner_rank, loser_rank, 
gs_num <- select(ungroup(gs_partial_pbp), winner_age,
                 w_pointswon, w_gameswon, w_setswon, l_pointswon, l_gameswon,
                 l_setswon, ave_serve_speed_p1,
                 ave_serve_speed_p2, n_aces_p1, n_aces_p2, n_winners_p1, n_winners_p2,
                 n_netpt_wpct_p1, n_netpt_wpct_p2, n_bp_wpct_p1, n_ue_p1, n_ue_p2,
                 n_sv_wpct_p1, n_sv_wpct_p2) %>% 
  na.omit()

mds <- cmdscale(dist(scale(gs_num)), k = 2) %>%
  as.data.frame()  # default column names will be V1, V2

mds$tournament <- select(ungroup(gs_partial_pbp), winner_age,
                 w_pointswon, w_gameswon, w_setswon, l_pointswon, l_gameswon,
                 l_setswon, ave_serve_speed_p1,
                 ave_serve_speed_p2, n_aces_p1, n_aces_p2, n_winners_p1, n_winners_p2,
                 n_netpt_wpct_p1, n_netpt_wpct_p2, n_bp_wpct_p1, n_ue_p1, n_ue_p2,
                 n_sv_wpct_p1, n_sv_wpct_p2, tournament) %>% 
  na.omit() %>% pull( tournament)

ggplot(mds, aes(V1, V2, label = tournament, color=tournament)) + geom_text()

gs_num <- select(ungroup(gs_partial_pbp), ave_serve_speed_p1,
                 ave_serve_speed_p2, n_aces_p1, n_aces_p2, n_winners_p1, n_winners_p2,
                 n_netpt_wpct_p1, n_netpt_wpct_p2, n_bp_wpct_p1, n_ue_p1, n_ue_p2,
                 n_sv_wpct_p1, n_sv_wpct_p2) %>% 
  na.omit()

mds <- cmdscale(dist(scale(gs_num)), k = 2) %>%
  as.data.frame()  # default column names will be V1, V2

mds$tournament <- select(ungroup(gs_partial_pbp), ave_serve_speed_p1,
                 ave_serve_speed_p2, n_aces_p1, n_aces_p2, n_winners_p1, n_winners_p2,
                 n_netpt_wpct_p1, n_netpt_wpct_p2, n_bp_wpct_p1, n_ue_p1, n_ue_p2,
                 n_sv_wpct_p1, n_sv_wpct_p2, tournament) %>% 
  na.omit() %>% pull( tournament)

ggplot(mds, aes(V1, V2, label = tournament, color=tournament)) + geom_text()

gs_num <- ungroup(filter(gs_partial_pbp, Tour == "atp")) %>%
    mutate(aces = n_aces_p1 + n_aces_p2) %>%
  select(n_winners_p1, n_winners_p2, n_ue_p1, n_ue_p2, aces) %>%
  na.omit()

mds <- cmdscale(scale(dist(gs_num)), k = 2) %>%
  as.data.frame()  # default column names will be V1, V2

mds$tournament <- ungroup(filter(gs_partial_pbp, Tour == "atp")) %>%
    mutate(aces = n_aces_p1 + n_aces_p2) %>%
  select(n_winners_p1, n_winners_p2, n_ue_p1, n_ue_p2, aces, tournament) %>% 
  na.omit() %>% pull( tournament)

mds2 <- bind_cols(mds, ungroup(filter(gs_partial_pbp, Tour == "atp")))

# mds$tournament[mds$tournament == "US Open"] <- "Australian Open"

ggplot(mds, aes(V1, V2, label = tournament, color=tournament)) + geom_point()

filter(mds2, V1 > 5)

filter(mds2, V2 > 2)

```

```