### 2019-02-21 contest/graphics/scratch.Rmd

---
title: "scratch"
author: "Kayla Frisoli, Shannon Gallagher, and Amanda Luby"
date: "September 15, 2018"
output: html_document
---

### Tall players and aces

Another common thought about tennis is that taller players are better servers. We find this to be true and see a strong positive correlation between teh variables. We highlight some interesting players BLAH BLAH BLAH.

```{r, fig.width=10, fig.height=4.8, fig.align='center'}

set.seed(8)

atp_servers <- c("Ivo Karlovic", "John Isner", "Rafael Nadal",
                 "Roger Federer", "Andy Murray")
atp_servers <- sapply(atp_servers,
                      function(x) return(sample(which(gs_players$name == x), 2))) %>%
  as.numeric() %>% gs_players[., ]

wta_servers <- c("Karolina Pliskova", "Caroline Wozniacki",
                 "Venus Williams", "Serena Williams", "Madison Keys")
wta_servers <- sapply(wta_servers,
                      function(x) return(sample(which(gs_players$name == x), 2))) %>%
  as.numeric() %>% gs_players[., ]

p_atp <- gs_players %>% filter(league == "ATP") %>%
ggplot( aes(log(ht+1), log(ace+1))) + 
  geom_jitter(aes(color = tournament)) + 
  scale_color_manual(" ", values = tournament_colors) + 
  geom_smooth( method = "lm", color="black") + 
  geom_label_repel(data=atp_servers,
            aes(log(ht+1), log(ace+1), label=name),
            alpha=.8, seed=1234, show.legend=FALSE) + 
  geom_label_repel(data=atp_servers,
            aes(log(ht+1), log(ace+1), label = name),
            seed=1234, fill=NA, show.legend=FALSE) +
  labs(x = "log(height + 1)", title = "ATP") +
  my_theme

p_wta <- gs_players %>% filter(league == "WTA") %>%
ggplot(aes(log(ht+1), log(ace+1))) + 
  geom_jitter(aes(color = tournament)) + 
  scale_color_manual(" ", values = tournament_colors) + 
  geom_smooth(method = "lm", color="black") + 
  geom_label_repel(data=wta_servers,
            aes(log(ht+1), log(ace+1), label=name),
            alpha=.8, seed=1234, show.legend=FALSE) + 
  geom_label_repel(data=wta_servers,
            aes( log(ht+1), log(ace+1), label=name),
            seed=1234, fill=NA, show.legend=FALSE) +
  labs(x = "log(height + 1)", title = "WTA")+
  my_theme

# grid.arrange(p_atp, p_wta, nrow=1, widths=c(.4, .6),
#              top = textGrob("Taller players are better servers: confirmed",
#                             gp=gpar(fontsize=20, fontfamily="serif")))

ggarrange(p_atp, p_wta, common.legend = TRUE, legend="bottom") %>%
  grid.arrange(top = textGrob("Taller players are better servers: confirmed",
                            gp=gpar(fontsize=20, fontfamily="serif")))

```